A community-maintained standard library of population genetic simulations

### 2019-12-05 popSimManu.tex
added mig rate plots and src

\documentclass[12pt,halfline,a4paper]{ouparticle}
% *
%
% ^.
\usepackage{amsmath,amssymb}
\usepackage{natbib}
\usepackage{graphicx,rotating}
\usepackage{dsfont}

@@

\newcommand{\stdpopsim}{\texttt{stdpopsim} }
\newcommand{\dadi}{$\partial a \partial i$ }
\newcommand{\MSMC}{\texttt{MSMC} }
\newcommand{\smcpp}{\texttt{smcpp} }
\newcommand{\stairwayplot}{\texttt{stairwayplot} }
\newcommand{\fastsimcoal}{\texttt{fastsimcoal} }

@@

\usepackage[colorinlistoftodos]{todonotes}

@@

\title{\stdpopsim: a community-supported toolkit for population genetic simulation}

@@

\abstract{The explosion in population genomic data demands ever more complex
modes of analysis, and increasingly these analyses depend on sophisticated simulations.
Recent advances in population genetic simulation have made it possible to
simulate large and complex models, but describing such models for a particular
simulation engine remains a difficult and error-prone task.
Computational genetics researchers currently re-implement simulation models
independently, leading to duplication of effort and reductions in reproducibility.
Population genetics, as a field, also lacks rigorous standard benchmarks by which new tools for inference might
be measured. Here we describe a new resource, stdpopsim, that attempts to rectify this situation.
Stdpopsim is a community-driven, open source project, which provides a standard
catalog of published simulation models from a wide range of organisms,
and supports multiple simulation engine backends.
% PopSim assures correctness in our implementations through a mature quality control pipeline that is community-based.
Here we describe the stdpopsim resource to encourange an even broader
community of developers to participate and to demonstrate some examples of how
it can be used to systematically compare demographic inference methods.}
\date{\today}

@@

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations, the distribution of fitness effects,
recombination rates, the extent of positive selection in genome sequence data
(\cite{li2011inference,schiffels2014inferring,terhorst2017robust,eyre2009estimating,kim2017inference,
chan2012genome,lin2013fast,Adrion662247,alachiotis2012omegaplus,degiorgio2016sweepfinder2,
kern2018diplos,sugden2018localization}).
While en masse these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
This is in large part because ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement as to what should
constitute community standards or best practices. The general modus operandi to date has been to
introduce methods which are validated by a set of bespoke simulations produced by individual
groups. Beyond a lack of efficiency from duplicated effort, such a situation leads
to decreases in reproducibility and transparency across the entire field.
Here we describe \texttt{stdpopsim}, a community-driven catalog
of standardized simulations across a variety of common study organisms
that will empower the field and usher in the next era of genetics data analysis.

Benchmarking the performance of inference methods in population genetics is not straightforward.
Inference methods often make different assumptions as
well as use different summaries of the genetic variation data, even when
attempting to address the same biological question. For example, there are
numerous methods to infer population size changes.
The \MSMC method (\cite{schiffels2014inferring}) allows for the
population to continuously change in size over time. Other methods, such as
\dadi (\cite{gutenkunst2009inferring}), allow for a smaller number of size changes to occur at certain
points throughout the population's history. Further, the types of data that the
methods use also differs. While \MSMC uses one to four whole genome sequences and leverages the
spatial patterns of genetic variation along the sequence, \dadi  instead uses the
site frequency spectrum (SFS), which summarizes the frequencies of SNPs in the
sample, ignoring any information about the correlation structure (linkage
disequilibrium) among SNPs. Thus from the start these methods may not be
expected to perform equally well in all scenarios. \adk{I think we are focussed
a bit too much on benchmarking currently in the intro. Could be good to add more on efficiency
and reproducibility}

The degree to which modeling assumptions and using different summaries of the
data can affect the ultimate results has been challenging to assess. Yet, there
have been several clear examples of different methods yielding fundamentally
different conclusions. For example, \MSMC methods applied to human genomes have
% JK is this right? > 100 years doesn't sound very ancient to me.
suggested large ancient ($>100$ years ago) ancestral population sizes and
bottlenecks that have not been detected by SFS-based methods (see
\cite{beichman2017comparison}). These distinct models differ in how they fit various summaries of the
genetic variation data, suggesting that modeling choices can greatly affect the
performance of the inference. Furthermore, it is reasonable
to assume generally that specific methods will perform better than others under certain scenarios.
However, because these conditions are typically unknown, empirical researchers
lack principled guidelines for deciding which statistical method is best suited
to accurately answer their particular question. The need for empirical
guidance will only increase as researchers continue to generate more genomic
variation data from non-model taxa and wish to make inferences regarding
demography and selection.

One way to benchmark different methods for statistical inference in population
genetics is to apply the methods to simulated datasets where the true model
parameters are known. Methods that perform well will accurately infer these
parameters and will also provide reasonable estimates of uncertainty of the
estimated parameters. Indeed, simulations studies are a standard ingredient in
most papers proposing new methods in population genetics. However, these
simulation studies are often specifically devised to showcase the performance
of the novel method and may not adequately compare performance to competing
methods. Moreover researchers are generally forced to ``reinvent the wheel''
with respect to simulations, and are forced to reimplement previously
inferred scenarios independently. This can lead to errors being incorporated
during the implementation of the simulations. Lastly, the degree to
which different methods are compared can be quite variable.

One reason why simulation studies do not compare the performance of many methods
is that these simulation studies are challenging to carry out. There are several
practical impediments to carrying out the ideal benchmarking simulation studies.
First, while there are many published population genetic models available, their
parameters are often not easily accessible, usually presented in supplemental
tables. Translating these parameters into input for a given simulation engine can
be a time-consuming and error prone process. In addition, papers often present many
different models, and different authors may choose to select distinct models,
making comparison difficult. While improvements in simulation software and ever
faster computers have partially ameliorating this concern, generating the
simulated datasets can still be a time and computing resource bottleneck.

For these reasons, we have generated a standardized, community-driven resource
for simulating  published demographic models from a number of popular study systems.
This resource, which we call \texttt{stdpopsim}, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community maintained catalog.
The \stdpopsim catalog currently contains four organisms: humans,
\emph{D.\ melanogaster}, \emph{A. thaliana}, and \emph{E. coli}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, the user can specify what
organism, map, chromosome, and demographic history they are interested in simulating, and are returned
simulation output from their chosen model.
The software has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and accuracy. Below we describe the resource and give an
example of how it can be used to benchmark demographic inference methods.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\linewidth]{display_items/Figure1.png}
\caption{\textbf{Structure of \stdpopsim}. \textbf{A.} The
hierarchical organization of the PopSim catalog contains all information
within individual species. Information within the species level is shown here
for \emph{Homo sapiens} only. We see that \emph{H. sapiens} is associated
with a default generation time and mutation rate, that it contains within
it a representation of the physical genome, genetic maps, and as shown a
subset of the demographic models currently implemented. \textbf{B.} Caching
and automated download of genetic maps. \textbf{C.} Example code to specify
and simulate models using \stdpopsim via the API (top) or
command-line interface (bottom). \textbf{D.} Simulation output from
\stdpopsim is generated as tree sequence files that can
be manipulated and analysed with \texttt{tskit}.}
\label{fig:cartoon}
\end{center}
\end{figure}

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution from PopSim currently is \stdpopsim, a community
maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{Drosophila},
\emph{Arabidopsis}, \emph{E. coli}, and \emph{Pongo}. A species definition consists
of two key elements.  Firstly, the library defines
some basic information about its genome, with information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically
observed heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of known maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the HapMapII (\cite{international2007second}) and
Decode (\cite{kong2010fine}) genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios.

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface (Fig. \ref{fig:cartoon}C), allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the “tree sequence”
format (\cite{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}), which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed very efficiently using the tskit library
(\cite{kelleher2016efficient,kelleher2018efficient}; Fig. \ref{fig:cartoon}D). Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator (\cite{kelleher2016efficient})
as the simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and have already implemented \texttt{SLiM} (\cite{haller2019slim}) as
an alternative back-end.

@@

\subsection*{\texttt{stdpopsim} Catalog}
The current contents of the \texttt{stdpopsim} catalog are shown in Table X. At the time of
writing the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Drosophila}, and \emph{Arabidopsis}. These include both
simpler, single population histories (e.g., in \emph{D. melanogaster} \cite{sheehan2016deep}),
and much more complex models which include population splitting, migration, and archaic
admixture (e.g., in humans \cite{ragsdale2019models}).

\begin{table}
\begin{footnotesize}
\input{model_table.tex}
\end{footnotesize}
\caption{\label{tab:catalog}
Details of the initial set of population models across four species.
\textbf{NOTES: Column names are ID, name (or description maybe?),
reference, CPU time, RAM}
}
\end{table}

@@

For non-human genomes we have have implemented two demographic histories for
\emph{Drosophila melanogaster} and one each from \emph{A. thaliana} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three epoch model estimated from
an African sample from \cite{sheehan2016deep} as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations. For For \emph{Arabidopsis thaliana}, we implemented the
model from in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes.

@@

We start by comparing popular methods for estimating
population size histories ($N(t)$) of single populations and subsequently
show simple examples of multi-population inference.
To evaluate and compare the performance of inference methods we developed
\texttt{snakemake} (\cite{koster2012snakemake}) workflows that allow reproducibility
in comparing inference methods which are available from \url{https://github.com/popgensims/analysis}
. The workflows allow efficient computing in multicore environments or on
cluster environments.

For single-population population size histories, we compared \MSMC, \smcpp, and
\stairwayplot (\cite{schiffels2014inferring,terhorst2017robust,liu2015exploring})
 on simulated genomes sampled from a single population,
in a number of the \stdpopsim models described above. The pipeline was generalized to
run $R$ replicates with $C$ chromosomes for $N$ samples, for a total of $R \times C$
simulations for each demographic model. After simulation, the pipeline prepares
input files for each of the respective inference methods by grouping all
chromosomes, for each sample, into a single file. This step results in an
input file for each simulation replicates $R$, pertaining to each of the
respective inference methods and derived from the same simulated tree sequences.
Each of the inference programs are then run in parallel, followed by plotting of
$N_e$ estimates from each program.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/homo_sapiens_mask_Ragsdale.png}
\caption{\textbf{Comparing estimates of $N(t)$ in humans}. Here we show estimates of population
size over time ($N(t)$) inferred using 4 different methods, \smcpp, \texttt{stairwayplot},
\MSMC with $n=2$ and $n=8$. Data were generated by simulating
replicate human genomes under the \cite{ragsdale2019models} model and using the genetic map
inferred in \cite{international2007second}. From top to bottom we show estimates for each
of the three populations in the model YRI, CEU, and CHB. In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_ragsdale}
\end{center}
\end{figure}

In Figure \ref{fig:n_t_ragsdale} we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under the \cite{ragsdale2019models} model using the
human genetic map estimated from \cite{international2007second}. On rows
we show inferred $N(t)$ for the three extant populations in the model.
On columns we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC). The solid black lines represent the true effective population
size as inferred from the inverse coalescent rate defined by the model.
Blue lines represent estimates from each of three replicate whole genome simulations.
While there is variation among methods and individual population estimates,
generally spoken the methods are accurate for this model of human history.

\stdpopsim allows us to readily compare these estimates to those drawn from a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but a different demographic
history-- that inferred by \cite{gutenkunst2009inferring}. Again we see that each
of the methods is capturing relevant parts of the population history, although the
degree of noise at any given time interval varies. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
sample.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/d_mel_Sheehan_mask2.png}
\caption{\textbf{Comparing estimates of $N(t)$ in \emph{Drosophila}}. Population
size over time ($N(t)$) estimated from an African population sample. Data were generated by simulating
replicate \emph{D.melanogaster} genomes under the three-epoch \cite{sheehan2016deep} model
with the genetic map inferred in \cite{comeron2012many}. In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_sheehan}
\end{center}
\end{figure}

As most population genetic methods development has been focused on human-scale
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from a demographic
model estimated from an African sample of \emph{Drosophila melanogaster} (\cite{sheehan2016deep}).
As with humans, we use \stdpopsim to simulate replicate \emph{Drosophila} genomes under
an empirically derived genetic map, and then try to infer back the parameters of model.
Accuracy is mixed among methods in this setting, and generally worse than what we
observe under simulations of the human genome.

\adk{do we want to include Arabidopsis results too?}

@@

In Figure \ref{fig:IM_popn_human}C we show estimates of population sizes and divergence
time from each of the methods for samples simulated for the African and European populations
under the \cite{gutenkunst2009inferring} model. Each of the methods compared has stronger
and weaker points to their accuracy across parameters. For instance the simple, IM-style
analyses we have performed do not capture recent population size change in CEU, however
they do a good job of estimating the quite simple YRI population size history.

@@

\section*{Discussion}
Here we have described what we view as the first product from the PopSim Consortium:
the \stdpopsim library. We have founded the Consortium with a number of specific aims
in mind: standardization of simulation within the computational genetics community,
increasing reproducibility of simulation based results, community-based development and
decision making in guiding best practices in population genetics, and eventual benchmarking
of inferential methods.

@@

We have shown in the results above how \stdpopsim can be used for direct comparisons
of inferential methods on a common set of simulations. While we have not made a concerted
effort at benchmarking methods yet, we believe that this would be an important effort for
the PopSim Consortium and population genetics more broadly. The analysis workflows that
we have released as part of \stdpopsim should provide a cookbook for the community to
dive in to more rigorous benchmarking than we have done thus far.

\adk{should we mention benchmarking of tree sequence inference here? what else?}

>>>>

\documentclass[12pt,halfline,a4paper]{ouparticle}
\usepackage{amsmath,amssymb}
\usepackage[round]{natbib}
\usepackage{graphicx,rotating}
\usepackage{dsfont}
\usepackage{xspace}

@@

\newcommand{\stdpopsim}{\texttt{stdpopsim}\xspace}
\newcommand{\dadi}{$\partial a \partial i$\xspace}
\newcommand{\MSMC}{\texttt{MSMC}\xspace}
\newcommand{\smcpp}{\texttt{smcpp}\xspace}
\newcommand{\stairwayplot}{\texttt{stairwayplot}\xspace}
\newcommand{\fastsimcoal}{\texttt{fastsimcoal}\xspace}

@@

\usepackage[hidelinks]{hyperref}

@@

\title{A community-maintained standard library of population genetic simulations}

@@

\abstract{The explosion in population genomic data demands ever more complex
modes of analysis, and increasingly these analyses depend on sophisticated simulations.
Recent advances in population genetic simulation have made it possible to
simulate large and complex models, but describing such models for a particular
simulation engine remains a difficult and error-prone task.
Computational genetics researchers currently re-implement simulation models
independently, leading to duplication of effort and the possibility for error.
Population genetics, as a field, also lacks standard benchmarks by which new tools for inference might
be measured. Here we describe a new resource, stdpopsim, that attempts to rectify this situation.
Stdpopsim is a community-driven, open source project, which provides a standard
catalog of published simulation models from a wide range of organisms,
and supports multiple simulation engine backends.
% PopSim assures correctness in our implementations through a mature, community-based quality control pipeline.
Here, we encourange an even broader
community of developers to participate and to demonstrate some examples of how
it can be used to systematically compare demographic inference methods.}
\date{\today}

@@

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations, the distribution of fitness effects,
recombination rates, the extent of positive selection in genome sequence data
\citep{li2011inference,schiffels2014inferring,terhorst2017robust,eyre2009estimating,kim2017inference,
chan2012genome,lin2013fast,Adrion662247,alachiotis2012omegaplus,degiorgio2016sweepfinder2,
kern2018diplos,sugden2018localization}.
While en masse these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming even more important.
This is in large part because ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement as to what should
constitute community standards or best practices. The general modus operandi to date has been to
introduce methods which are validated by a set of bespoke simulations produced by individual
groups. Beyond a lack of efficiency from duplicated effort, such a situation leads
to decreases in reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized simulations across a variety of common study organisms
that will empower the field and usher in the next era of genetics data analysis.

The degree to which modeling assumptions and different summaries of the
data can affect the ultimate results has been challenging to assess. Yet, there
have been several clear examples of different methods yielding fundamentally
different conclusions. For example, \MSMC methods applied to human genomes have
% JK is this right? > 100 years doesn't sound very ancient to me.
suggested large ancient ($>100$ years ago) ancestral population sizes and
bottlenecks that have not been detected by SFS-based methods
\citep[see][]{beichman2017comparison}.
These distinct models differ in how they fit various summaries of the
genetic variation data, suggesting that modeling choices can greatly affect the
performance of the inference. Furthermore, it is reasonable
to assume generally that specific methods will perform better than others under certain scenarios.
However, because these conditions are typically unknown, empirical researchers
lack principled guidelines for deciding which statistical method is best suited
to accurately answer their particular question. The need for empirical
guidance will only increase as researchers continue to generate more genomic
variation data from non-model taxa and wish to make inferences regarding
demography and selection.

One way to benchmark different methods for statistical inference in population
genetics is to apply the methods to simulated datasets where the true model
parameters are known. Methods that perform well will accurately infer these
parameters and will also provide reasonable estimates of uncertainty of the
estimated parameters. Indeed, simulations studies are a standard ingredient in
most papers proposing new methods in population genetics. However, these
simulation studies often use simulations of the particular models for which the method was designed,
which showcases the novel method rather than rigorously comparing performance to competing methods.
Of course, testing a new method in the situation for which it was designed is an important step,
but more rigorous testing is necessary for tools that are to be widely used in the field.
% Moreover researchers are generally forced to reimplement previously
% inferred scenarios independently, which makes good benchmarking
% a major additional time investment, and can lead to errors.
% Lastly, the degree to which different methods are compared can be quite variable.
Many more researchers in the broader field
use simulations to perform power and other exploratory analyses.
The many steps of implementing a moderately complicated population genetics simulation,
including translating published demographic models to a simulator
and obtaining genetic maps,
present a major obstacle, especially to new researchers.

% One reason why simulation studies do not compare the performance of many methods
% across a wide range of situations
% is that simulation studies are challenging to carry out, for several reasons.
% First, while there are many published population genetic models available, their
% parameters are often not easily accessible, usually presented in supplemental
% tables. Translating these parameters into input for a given simulation engine can
% be a time-consuming and error prone process. In addition, papers often present many
% different models, and different authors may choose to select distinct models,
% making comparison difficult. While improvements in simulation software and ever
% faster computers have partially ameliorated this concern, generating the
% simulated datasets can still be a time and computing resource bottleneck.

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \texttt{stdpopsim}, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community maintained catalog.
The \stdpopsim catalog currently contains four organisms: humans,
\emph{D.~melanogaster}, \emph{A.~thaliana}, and \emph{E.~coli}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, the user can specify what
organism, genetic map, chromosome, and demographic history they are interested in simulating, and are returned
simulation output from their chosen model.

The software has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and accuracy.
Importantly, we have rigorous quality control methods to ensure implemented models are accurate,
and have documented methods for others to contribute new modules.
We welcome new collaborators.
Below we describe the resource and give an
example of how it can be used to benchmark demographic inference methods.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\linewidth]{display_items/Figure1.pdf}
\caption{\textbf{Structure of \stdpopsim}. \textbf{A.} The
hierarchical organization of the PopSim catalog contains all information
within individual species. Information within the species level is shown here
for \emph{Homo sapiens} only. We see that \emph{H. sapiens} is associated
with a default generation time and mutation rate, that it contains within
it a representation of the physical genome, genetic maps, and as shown a
subset of the demographic models currently implemented. \textbf{B.} Caching
and automated download of genetic maps. \textbf{C.} Example code to specify
and simulate models using \stdpopsim via the API (top) or
command-line interface (bottom). \textbf{D.} Simulation output from
\stdpopsim is generated as tree sequence files that can
be manipulated and analysed with \texttt{tskit}.}
\label{fig:cartoon}
\end{center}
\end{figure}

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution from PopSim currently is \stdpopsim, a community
maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{Drosophila},
\emph{Arabidopsis}, \emph{E. coli}, and \emph{Pongo}. A species definition consists
of two key elements.  Firstly, the library defines
some basic information about its genome, with information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically
observed heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of known maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the
HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios.

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface (Fig. \ref{fig:cartoon}C), allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the “tree sequence”
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed very efficiently using the tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and have already implemented \texttt{SLiM} \citep{haller2019slim} as
an alternative back-end.

@@

\subsection*{\texttt{stdpopsim} Catalog}
The current contents of the \texttt{stdpopsim} catalog are shown in Table X. At the time of
writing the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Drosophila}, and \emph{Arabidopsis}. These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.

\begin{table}[t]
\begin{footnotesize}
\input{model_table.tex}
\end{footnotesize}
\caption{\label{tab:catalog}
Details of the initial set of population models across four species.
\textbf{NOTES: Column names are ID, name (or description maybe?),
reference, CPU time, RAM}
}
\end{table}

@@

For non-human genomes we have have implemented two demographic histories for
\emph{Drosophila melanogaster} and one each from \emph{A. thaliana} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three epoch model estimated from
an African sample from \cite{sheehan2016deep} as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations. For \emph{Arabidopsis thaliana}, we implemented the
model from in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes.

@@

We start by comparing popular methods for estimating
population size histories ($N(t)$) of single populations and subsequently
show simple examples of multi-population inference.
To evaluate and compare the performance of inference methods we developed
\texttt{snakemake} \citep{koster2012snakemake} workflows that allow reproducibility
in comparing inference methods which are available from \url{https://github.com/popgensims/analysis}
. The workflows allow efficient computing in multicore environments or on
cluster environments.

For single-population population size histories, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the \stdpopsim models described above. The pipeline was generalized to
run $R$ replicates with $C$ chromosomes for $N$ samples, for a total of $R \times C$
simulations for each demographic model. After simulation, the pipeline prepares
input files for each of the respective inference methods by grouping all
chromosomes, for each sample, into a single file. This step results in an
input file for each simulation replicates $R$, pertaining to each of the
respective inference methods and derived from the same simulated tree sequences.
Each of the inference programs are then run in parallel, followed by plotting of
$N_e$ estimates from each program.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/homo_sapiens_mask_Ragsdale.pdf}
\caption{\textbf{Comparing estimates of $N(t)$ in humans}. Here we show estimates of population
size over time ($N(t)$) inferred using 4 different methods, \smcpp, \texttt{stairwayplot},
\MSMC with $n=2$ and $n=8$. Data were generated by simulating
replicate human genomes under the \cite{ragsdale2019models} model and using the
HapMapII genetic map~\citep{international2007second}. From top to bottom we show estimates for each
of the three populations in the model YRI, CEU, and CHB. In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_ragsdale}
\end{center}
\end{figure}

% JK here --- I think it would really help if we described what the models
% were, briefly, rather than saying who inferred them.
In Figure \ref{fig:n_t_ragsdale} we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under a sophisticated demographic model
and empirical genetic map. On rows
we show inferred $N(t)$ for the three extant populations in the model.
On columns we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC). The solid black lines represent the true effective population
size as inferred from the inverse coalescent rate defined by the model.
Blue lines represent estimates from each of three replicate whole genome simulations.
While there is variation among methods and individual population estimates,
generally spoken the methods are accurate for this model of human history.

\stdpopsim allows us to readily compare these estimates to those drawn from a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but a different demographic
history -- that inferred by \cite{gutenkunst2009inferring}. Again we see that each
of the methods is capturing relevant parts of the population history, although the
discrepancy between inferred and simulated values is changes over time. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

We can even do away with the demographic model entirely, and simulate under a human
genome architecture and genetic map to see how well methods might to at recovering
the population history of a constant sized population. We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}. Again we see that all methods
are recovering a modern population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to learn sinusoidal
patterns of population size even though no change is present.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/d_mel_Sheehan_mask2.pdf}
\caption{\textbf{Comparing estimates of $N(t)$ in \emph{Drosophila}}. Population
size over time ($N(t)$) estimated from an African population sample. Data were generated by simulating
replicate \emph{D.melanogaster} genomes under the three-epoch \cite{sheehan2016deep} model
with the genetic map inferred in \cite{comeron2012many}. In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_sheehan}
\end{center}
\end{figure}

As most population genetic methods development has been focused on human-scale
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from a demographic
model estimated from an African sample of \emph{Drosophila melanogaster} \citep{sheehan2016deep}.
As with humans, we use \stdpopsim to simulate replicate \emph{Drosophila} genomes under
an empirically derived genetic map, and then try to infer back the parameters of model.
Accuracy is mixed among methods in this setting, and generally worse than what we
observe under simulations of the human genome.

@@

In Figure \ref{fig:IM_popn_human}C we show estimates of population sizes and divergence
time from each of the methods for samples simulated for the African and European populations
under the \cite{gutenkunst2009inferring} model. Each of the methods compared has stronger
and weaker points to their accuracy across parameters. For instance the simple, IM-style
analyses we have performed do not capture recent population size change in CEU, however
they do a good job of estimating the simpler simulated YRI population size history.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}

Here we have described what we view as the first product from the PopSim Consortium:
the \stdpopsim library. We have founded the Consortium with a number of specific aims
in mind: standardization of simulation within the computational genetics community,
increasing reproducibility of simulation based results, community-based development and
decision making in guiding best practices in population genetics, and eventual benchmarking
of inferential methods.

@@

We have shown in the results above how \stdpopsim can be used for direct comparisons
of inferential methods on a common set of simulations. While we have not made a concerted
effort at benchmarking methods yet, we believe that this would be an important goal for
the PopSim Consortium and population genetics more broadly. The analysis workflows that
we have released as part of \stdpopsim should provide a cookbook for the community to
dive in to more rigorous benchmarking than we have done thus far.

In the use case comparisons we have performed,
we found that the methods tested \emph{do} generally perform well on human genomes,
up to, for instance, the usual factor-of-two differences in inferred population sizes.
This will not be surprising to people familiar with the landscape of
demographic inference methods in population genetics,
but the nature and scale of uncertainty may be not clear to people new to the field.
The ability for \stdpopsim to generate data with and without significant features, such
as the genetic map or population size change (e.g., Figure \ref{fig:n_t_HomSap_constant}), allows
us to ask directly about the various failure modes of popular methods.
Moreover the comparison of methods across the various genome organizations, genetic maps,
and demographic histories of different organisms provides valuable information
about how methods might perform on non-human systems.
In our initial comparison we find that inference of population histories from \emph{Drosophila}
or \emph{Arabidopsis} simulated genomes is a fair bit more difficult from SMC based methods
than in the human genome. Likely this is consequence of the reduced physical and genetic
map lengths of these genomes, although we will leave further investigation of this to the future.
Finally, comparison of results across methods or simulation runs
gives a strong estimate of inference uncertainty, analogously to parametric bootstrapping --
although different methods have different strengths and weaknesses,
strong conclusions should probably not rest entirely on features inferred by only one method.

The broad brush comparisons we have done, while far from exhaustive, are informative,
and highlight the utility of easy comparison of methods on the same footing.
The tests here were under best-case scenarios --
simulations of discrete, randomly mating populations that fit the assumptions of the methods nearly exactly --
and yet different methods, as well as applications of the same method
to different simulations of the same model,
produced somewhat different results.

%\adk{should we mention benchmarking of tree sequence inference here? what else?}
